
\hide{
\todo{Bo-Yin, please describe how the ladder step is used and why it is 
important for about .5 page. Maybe another .5 page for its mathematical 
properties? }
}

Recall that X25519 is based on the elliptic curve Curve25519. 
Cryptographic primitives of the key exchange protocol perform
sequences of algebraic operations on the Abelian group $\bbfG$ induced
by Curve25519, not on the finite field $\bbfGF(\varrho)$. As
aforementioned, the binary operation $\Gplus$ requires another
sequence of arithmetic computation over $\bbfGF(\varrho)$. Errors
could still be introduced or even implanted in implementations of
$\Gplus$. Correctness of arithmetic constructs over $\bbfGF(\varrho)$
does not necessarily imply the correctness of the mathematical
construct in X25519. The Montgomery Ladderstep is the mathematical
construct widely used to implement $\Gplus$ on Curve25519. We 
verify a low-level implementation of the mathematical construct in
this experiment. 

\begin{algorithm}[h]
\label{evaluation:ladder-step:montgomery}
\begin{algorithmic}[1]
\Function{ladderstep}{$x_1, x_m, z_m, x_n, z_n$}
\begin{multicols}{3}
\State $t_1 \leftarrow x_m \Fplus z_m$
\State $t_2 \leftarrow x_m \Fminus z_m$
\State $t_7 \leftarrow t_2 \Ftimes t_2$
\State $t_6 \leftarrow t_1 \Ftimes t_1$
\State $t_5 \leftarrow t_6 \Fminus t_7$
\State $t_3 \leftarrow x_n \Fplus z_n$
\State $t_4 \leftarrow x_n \Fminus z_n$\rule{0ex}{0ex}
\State $t_9 \leftarrow t_3 \Ftimes t_2$
\State $t_8 \leftarrow t_4 \Ftimes t_1$
\State $x_n \leftarrow t_8 \Fplus t_9$
\State $z_n \leftarrow t_8 \Fminus t_9$
\State $x_n \leftarrow x_{n} \Ftimes x_{n}$
\State $z_n \leftarrow z_{n} \Ftimes z_{n}$
\State $z_n \leftarrow z_{n} \Ftimes x_1$\rule{0ex}{0ex} 
\State $x_m \leftarrow t_6 \Ftimes t_7$
\State $z_m \leftarrow 121666 \Ftimes t_5$
\State $z_m \leftarrow z_m \Fplus t_7$
\State $z_m \leftarrow z_m \Ftimes t_5$
\State \Return $(x_m, z_m, x_n, z_n)$
\EndFunction
\end{multicols}
\end{algorithmic}
\caption{Montgomery Ladderstep}
\end{algorithm}

Let $p \in G$ be an element of the group $\bbfG$ and hence a point on
the elliptic curve Curve25519. We write $n \cdot p$ for
$\overbrace{p \Gplus \cdots \Gplus p}^n \in G$ when $n \in \bbfN$. A
crucial mathematical construct in X25519 is to compute the
$x$-coordinate of $n \cdot p$ for $n \in \bbfN$. The na\"ive
implementation would follow the equation (\ref{eq:curve25519-group})
and compute both $x$- and $y$-coordinates with, in particular, a
division operation $\Fdiv$ in $\bbfGF(\varrho)$. In~\cite{M:87:SPEC},
a more efficient implementation is proposed. 

$np = x_n \Fdiv z_n$.
$\textsc{ladderstep} (x_1, x_m, z_m, x_n, z_n) = (x_{2m}, z_{2m},
x_{m+n}, z_{m+n})$
 
\todo{mention the coefficient $4$ in the ratio}